---
title: "Lab 03 - Nobel laureates"
author: "Ben Hardin"
date: "1-23-2023"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

There are 26 variables and 935 observations. In this case, the observations are people (nobel prize winning people!)

```{r num-columns-rows}
glimpse(nobel)
```

### Exercise 2

This code makes a new dataframe which contains only nobel laureates who (1) have country data available, (2) are people and not organizations, & (3) are still alive. By glimpsing at the data, we confirm that the new dataframe contains 228 people.

```{r make-new-df}
nobel_living <- nobel %>%
  filter(
    !is.na(country),
    gender != "org",
    is.na(died_date))

#verifying this code worked
glimpse(nobel_living)
```
### Exercise 3

```{r getting-usa-scientists}
#creating USA variable
nobel_living <- nobel_living %>%
  mutate(country_us = if_else(country == "USA", "USA", "other"))

#filtering for sciences
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

```{r visualizing-category-country}
ggplot(data = nobel_living_science,
       aes(x = country_us,
           fill = country_us))+
  geom_bar(color = "black")+
  facet_wrap(~ category)+
  coord_flip()+
  labs(y = "Number of Nobel Prizes",
       x = "Country at time of award",
       color = "Country")
```
### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
