---
title: "Lab 03 - Nobel laureates"
author: "Ben Hardin"
date: "1-23-2023"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

There are 26 variables and 935 observations. In this case, the observations are people (nobel prize winning people!)

```{r num-columns-rows}
glimpse(nobel)
```

### Exercise 2

This code makes a new dataframe which contains only nobel laureates who (1) have country data available, (2) are people and not organizations, & (3) are still alive. By glimpsing at the data, we confirm that the new dataframe contains 228 people.

```{r make-new-df}
nobel_living <- nobel %>%
  filter(
    !is.na(country),
    gender != "org",
    is.na(died_date))

#verifying this code worked
glimpse(nobel_living)
```
### Exercise 3

This code chunk creates the dataset that we need to answer the question in Exercise 3, showing how many nobel laureates in the sciences were based in the US or somewhere else when they got their prize.

```{r getting-usa-scientists}
#creating USA variable
nobel_living <- nobel_living %>%
  mutate(country_us = if_else(country == "USA", "USA", "other"))

#filtering for sciences
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

The visualization shows that, generally, nobel prizes for the sciences have more often been awarded to people living in the US than people living elsewhere. This difference is the greatest for Economics, while the difference is smallest for Chemistry. This supports Buzzfeed's claim that "Most living Nobel laureates in the sciences are based in the US."

```{r visualizing-category-country}
ggplot(data = nobel_living_science,
       aes(x = country_us))+
  geom_bar(color = "black")+
  facet_wrap(~ category)+
  coord_flip()+
  labs(y = "Number of Nobel Prizes",
       x = "Country at time of award")
```
### Exercise 4

The code below creates a new variable that indicates whether a nobel laureate was born in the US or in another country. 105 Nobel prize winners in the sciences were born in the US.

```{r make-born-country}
nobel_living_science <- nobel_living_science %>%
  mutate(born_country_us = if_else(born_country == "USA", "USA", "other"))

#getting born in the usa count
nobel_living_science %>%
  count(born_country_us)
```
### Exercise 5

The bar plots below show that, of those Nobel laureates who were living in the US when they received their award, about 1/3 to 1/6 of them (depending on the specific field) were born in another country. This could either support or not support Buzzfeed's claim that "many [US-based Nobel laureates] were born in other countries", depending on how "many" is defined. It is true based on this graph that Nobel laureates are much more likely to have immigrated to the US from another country than vice versa, and it is true that a sizeable chunk of American Nobel laureates are immigrants. On the other hand, if we misunderstood Buzzfeed's claim to be that MOST American Nobel laureates immigrated to the US from other countries, then these data would not support that claim.

```{r visualizing-country-by-birth-country}
ggplot(data = nobel_living_science,
       aes(x = country_us,
           fill = born_country_us))+
  geom_bar(color = "black", position = "stack")+
  scale_fill_manual(values = c("grey", "black"))+
  facet_wrap(~ category)+
  coord_flip()+
  labs(y = "Number of Nobel Prizes",
       x = "Country at time of award",
       fill = "Country of birth")
```


### Exercise 6

This code shows us where those American nobel laureates who immigrated from other countries were originally born. The two most common countries are Germany and the UK.

```{r getting-count-of-origin-countries}
nobel_living_science %>%
  filter(country_us == "USA", born_country_us == "other") %>%
  count(born_country) %>%
  arrange(desc(n))
```